---
title: "Editors' Guide"
author: "First draft by Di Cook"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Operating procedures for the editors}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = FALSE, 
  warning = FALSE,
  message = FALSE
)
```

# Mission

Executive Editor (EE) of the R Journal are the primary article handler for about a quarter of submissions per year, responsible for:

- matching the article with an appropriate Associate Editor (AE) or handling obtaining reviews without an AE.
- communicating with AEs on submitted articles, and if necessary providing guidance to an AE about use of how GitHub is used, and relevant `rj` functionality.
- communicating progress and reviews with the corresponding author.
- making a final decision if an article should be accepted, or rejected, or on an appropriate revision status, usually based on reviewer and AE feedback. 
- maintaining and enhancing the `rj` and `rjtools` software, in collaboration with other EEs and AEs.

# Overview of editorial board

The editorial board consists of four editors, including an Editor-in-Chief (EIC), and a non-fixed number of associate editors. Each editor serves a roughly four year term. Associate editors serve a three year term, generally beginning in the calendar year of their invitation, and ending at the end of their third year.

In the third calendar year of their term the EE acts as EIC. The EIC takes primary responsibility for:

- Production of four journal issues per year, in March, June, September, and December. 
  The issue typically appears a month or two after the publication date, because articles accepted up to June 30, and mid-Dec, respectively will likely be included. 
- Managing the flow of submissions, ensuring that these are handled in a timely manner.
- Maintaining the R Journal web site http://journal.r-project.org. 
- Maintaining the formatting and tools used to manage the journal.
- Ensuring smooth communication between editors and associate editors and authors.

EE's have the primary responsibility for handling any particular submission.

AE's have the primary responsibility for obtaining reviews and making recommendations for their assigned articles. 

# New editors and associate editors

New editors and associate editors should be added to the `editors.csv` and `associate_editors.csv` files in `inst` folder. 

# Article assignment to AE

An article repo needs to be created for any new AEs in the for of the articles repo, named `ae-articles-XX`. Initials replace XX and match the codes in the `csv` file.

The AE should be notified of any new assignment.

# Communications

A newly assigned article for an AE needs to have the directory need to be copied from the `articles` repo to the relevant `ae-articles-XX`. The `articles` repo is the master copy, and once a paper is retrieved from an AE, the `DESCRIPTION` file is updated, and the `correspondence` folder is populated. 

Slack is used for communication, between EIC, EEs and AEs, and general information about operations. The channel `editors_private` can be used for protected conversations between EEs. 

Weekly virtual meetings are held to keep communication between editors current.

Email is the primary manner for communicating with authors. The EE, not AEs, should communicate with authors. 

# Workflow and operations

## Getting started

Install the `rj` package with 

```{r eval=FALSE, echo=TRUE}
remotes::install_github("rjournal/rj")
```

The package is being updated and revised regularly, so you may want to re-install occasionally.

## Potential AE

The keywords of a paper should be matched against the keywords of AEs available in `associate_editors.csv`. Also, the function `ae_workload()` should be used to ensure that the potential AE hasn't got too many current assignments.

## Timeline for reviews

This is what the acknowledgement letter tells authors about the timeline for their paper to be handled:

```
The submission process proceeds as follows:

* review by editorial board & assignment of associate editor (~1-2 weeks)
* reviewers solicited (~1-2 weeks)
* reviews received (~1-2 months)
```


## Workflow

```{r wfgraph, fig.width=8, fig.height=12, fig.align='center'}
library(DiagrammeR)
editor_df <- create_node_df(n=11,  
    label=c("is this a new submission?", # 1
            "examine article: should the paper progress to review?", # 2
            "is there a suitable AE?", # 3
            "update status to be `with AE`, update_status(), copy folder to AEs repo", # 4
            "message/email AE", # 5
            "has AE returned reviews and decision?", # 6
            "add_review() to correspondence folder, send decision to author, \n with reviews if appropriate, update_status(), \n notify and thank AE about decision, if appropriate", # 7
            "is the revision adequate?", # 8
            "should it go back to AE?", # 9
            "reject(), contact authors", # 10
            "act as AE, find reviewers, update status, follow AE flow chart"), # 11
   shape = c("ellipse", "ellipse", "ellipse", 
             "rectangle", "rectangle", 
             "ellipse", "rectangle", "ellipse", 
             "ellipse", "rectangle", "rectangle"),
   width = c(3, 5, 3, 5, 2, 3, 4, 3, 3, 3, 4),
   fillcolor = c("orange", "orange", "orange", 
                 "grey90", "grey90",
                 "orange", "salmon", "orange", 
                 "orange", "salmon", "turquoise"),
   fontcolor = "black")

editor_edge_df <- create_edge_df(
  from = c(1,2,3,4,5,6,1,8, 8,9,9, 2, 3, 6), 
  to =   c(2,3,4,5,6,7,8,9,10,7,4,10, 11, 5), 
  label = c("Y", "Y", "Y", ".", ".", "Y",
            "N","Y","N", "N", "Y", "N", "N", "N"),
  color = c("dodgerblue2", "dodgerblue2", "dodgerblue2", 
            "grey50", "grey50", "dodgerblue2",
            "red", "dodgerblue2", "red", "red", 
            "dodgerblue2", "red", "red", "red"))

editor <- create_graph(editor_df, 
                       editor_edge_df, 
                       directed = TRUE, attr_theme = "tb")
render_graph(editor)
```

## Dealing with problematic authors

`vexing_authors.csv`
