---
title: "Editors' Guide"
author: "First draft by Di Cook"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Operating procedures for the editors}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = FALSE, 
  warning = FALSE,
  message = FALSE
)
```

# Mission

As an Executive Editor (EE) for the R Journal you will be the primary article handler for about a quarter of submissions per year, with  assignments given by the Editor-in-Chief (EIC). You will be responsible for:

- matching the article with an appropriate Associate Editor (AE) or handling the reviews yourself.
- communicating with AEs on submitted articles, and if necessary providing guidance to an AE about use of how GitHub is used, and relevant `rj` functionality
- communicating progress and reviews with the corresponding author.
- deciding if an article should be accepted, or rejected, or on an appropriate revision status, usually based on reviewer and AE feedback. 
- maintaining and enhancing the `rj` and `rjtools` software, in collaboration with other editors.

# Editorial structure

The editorial board consists of four editors, including an Editor-in-Chief, and a non-fixed number of associate editors. Each editor serves a term of four years, starting on 1 January and ending on 31 December four years later. Associate editors serve a three year term, generally beginning in the calendar year of their invitation, and ending at the end of their third year.

In the third year of their term the editor acts as Editor-in-Chief. The EIC takes primary responsibility for:
- Production of two journal issues per year, in June and December.
- Managing the flow of submissions, ensuring that these are handled in a timely manner.
- Maintaining the R Journal web site http://journal.r-project.org
- Maintaining the formatting and tools used to manage the journal.
- Ensuring smooth communication between editors and associate editors.

# New editors and associate editors

New editors and associate editors should be added to the `editors.csv` and `associate_editors.csv` files in `inst` folder.

# Communications

Article repos need to be copied to the relevant `ae-articles-XX`. The `articles` repo is the master copy, and once a paper is retrieved from an AE, the `DESCRIPTION` file is updated, and the `correspondence` folder is populated. 

Slack is used for communication, between EIC, EEs and AEs, and general information about operations. The channel `editors_private` can be used for protected conversations between EEs. 

Weekly virtual meetings are held to keep communication between editors current.

# Workflow and operations

## Getting started

Install the `rj` package with 

```{r eval=FALSE, echo=TRUE}
remotes::install_github("rjournal/rj")
```

The package is being updated and revised regularly, so you may want to re-install occasionally.

## Potential AE

The keywords of a paper should be matched against the keywords of AEs available in `associate_editors.csv`. Also, the function `ae_workload()` should be used to ensure that the potential AE hasn't got too many current assignments.

## Timeline for reviews

add note from template letter to authors

## Workflow

```{r wfgraph, fig.width=8, fig.height=12, fig.align='center'}
library(DiagrammeR)
editor_df <- create_node_df(n=14,  
    label=c("check status of papers being handled: summarise_articles()", # 1
            "has the paper been acknowledged?", # 2
            "acknowledge paper \n acknowledge_submission() ", # 3
            "is this a new submission?", # 4
            "examine article: should the paper progress to review?", # 5
            "is there a suitable AE?", # 6
            "update status to be `with AE`, update_status(), copy repo to AEs", # 7
            "message/email AE, hopefully they accept assignment", # 8
            "has AE returned reviews and decision?", # 9
            "send reviews and decision to author, update_status(), \n notify/thank AE about decision", # 10
            "is the revision adequate?", # 11
            "should it go back to AE?", # 12
            "reject, update status(), contact authors", # 13
            "act as AE, find reviewers, update status, follow AE flow chart"), # 14
   shape = c("rectangle", "ellipse", "rectangle", 
             "ellipse", "ellipse", "ellipse", 
             "rectangle", "rectangle", 
             "ellipse", "rectangle", "ellipse", 
             "ellipse", "rectangle", "rectangle"),
   width = c(5, 3, 2, 3, 7, 3, 5, 5, 3, 4, 3, 3, 3, 4),
   fillcolor = c("grey90", "orange", "grey90",  
                 "orange", "orange", "orange", 
                 "grey90", "grey90",
                 "orange", "salmon", "orange", 
                 "orange", "salmon", "turquoise"),
   fontcolor = "black")

editor_edge_df <- create_edge_df(
  from = c(1,2,2,4,5,6,7,8,9,3,4,11,12,12,5,6,9), 
  to = c(2,3,4,5,6,7,8,9,10,4,11,12,7,13,13,14,8), 
  label = c(".", "Y", ".", "Y", "Y", "Y", ".", ".",
            "Y",".","N", "Y", "Y", "N", "N","N"))

editor <- create_graph(editor_df, 
                       editor_edge_df, 
                       directed = TRUE, attr_theme = "tb")
render_graph(editor)
```

## Dealing with problematic authors

`vexing_authors.csv`
