---
title: "Editors' Guide"
author: "Di Cook and Rob J Hyndman"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Operating procedures for the editors}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)
```

# Mission

Each Executive Editor (EE) of the R Journal is the primary article handler for about a quarter of submissions per year, responsible for:

- matching the article with an appropriate Associate Editor (AE) or handling obtaining reviews without an AE.
- communicating with AEs on submitted articles, and if necessary providing guidance to an AE about use of how GitHub is used, and relevant `rj` functionality.
- communicating progress and reviews with the corresponding author.
- making a final decision if an article should be accepted, or rejected, or on an appropriate revision status, usually based on reviewer and AE feedback.
- maintaining and enhancing the `rj` and `rjtools` software, in collaboration with other EEs and AEs.

# Article assignment to AE

An article repo needs to be created for any new AEs in the for of the articles repo, named `ae-articles-XX`. Initials replace XX and match the codes in the `csv` file.

The AE should be notified of any new assignment.

# Communications

A newly assigned article for an AE needs to have the directory copied from the `articles` repo to the relevant `ae-articles-XX`. The `articles` repo is the master copy, and once a paper is retrieved from an AE, the `DESCRIPTION` file is updated, and the `correspondence` folder is populated.

Slack is used for communication, between EIC, EEs and AEs, and general information about operations. The channel `editors_private` can be used for protected conversations between EEs.

Monthly virtual meetings are held to keep communication between editors current.

Email is the primary manner for communicating with authors. The EE, not AEs, should communicate with authors.

# Workflow and operations

## Getting started

Install the `rj` package with

```{r eval=FALSE, echo=TRUE}
remotes::install_github("rjournal/rj")
```

The package is being updated and revised regularly, so you may want to re-install occasionally.

## Potential AE

The keywords of a paper should be matched against the keywords of AEs available in `associate_editors.csv`. Also, the function `ae_workload()` should be used to ensure that the potential AE hasn't got too many current assignments.

## Timeline for reviews

This is what the acknowledgement letter tells authors about the timeline for their paper to be handled:

```
The submission process proceeds as follows:

* review by editorial board & assignment of associate editor (~1-2 weeks)
* reviewers solicited (~1-2 weeks)
* reviews received (~1-2 months)
```


## Workflow

```{r wfgraph, fig.width=8, fig.height=12, fig.align='center'}
library(DiagrammeR)
editor_df <- create_node_df(n=11,
    label=c("is this a new submission?", # 1
            "examine article: should the paper progress to review?", # 2
            "is there a suitable AE?", # 3
            "update status to be `with AE`, update_status(), copy folder to AEs repo", # 4
            "message/email AE", # 5
            "has AE returned reviews and decision?", # 6
            "add_review() to correspondence folder, send decision to author, \n with reviews if appropriate, update_status(), \n notify and thank AE about decision, if appropriate", # 7
            "is the revision adequate?", # 8
            "should it go back to AE?", # 9
            "reject(), contact authors", # 10
            "act as AE, find reviewers, update status, follow AE flow chart"), # 11
   shape = c("ellipse", "ellipse", "ellipse",
             "rectangle", "rectangle",
             "ellipse", "rectangle", "ellipse",
             "ellipse", "rectangle", "rectangle"),
   width = c(3, 5, 3, 5, 2, 3, 4, 3, 3, 3, 4),
   fillcolor = c("orange", "orange", "orange",
                 "grey90", "grey90",
                 "orange", "salmon", "orange",
                 "orange", "salmon", "turquoise"),
   fontcolor = "black")

editor_edge_df <- create_edge_df(
  from = c(1,2,3,4,5,6,1,8, 8,9,9, 2, 3, 6),
  to =   c(2,3,4,5,6,7,8,9,10,7,4,10, 11, 5),
  label = c("Y", "Y", "Y", ".", ".", "Y",
            "N","Y","N", "N", "Y", "N", "N", "N"),
  color = c("dodgerblue2", "dodgerblue2", "dodgerblue2",
            "grey50", "grey50", "dodgerblue2",
            "red", "dodgerblue2", "red", "red",
            "dodgerblue2", "red", "red", "red"))

editor <- create_graph(editor_df,
                       editor_edge_df,
                       directed = TRUE, attr_theme = "tb")
render_graph(editor)
```

## Dealing with problematic authors

A few problematic authors are listed in `vexing_authors.csv`

Please advise the EIC if you think anyone needs to be added to this list.
