---
title: "Editor-in-Chief's Guide"
date: "Last updated: 2 January 2026"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Guide for the Editor-in-Chief}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)
```

# Downloading new submissions and revisions

Articles are submitted through a [Google form](https://forms.gle/ykj6QcoGQD5ctnA5A) which populates a [Google sheet](https://docs.google.com/spreadsheets/d/15Tem82ikjHhNVccZGrHFVdzVjoxAmwaiNwaBE95wu6k/edit#gid=1671813370), and a [Google drive](https://drive.google.com/drive/folders/0B2aFIue-Ar8dfmdVdHlpTnRndWhxQnNtR0o2YnoycDU1cVVzRlRaM0VSQWU1YUlnZlkyZEk?resourcekey=0-pzsHNE5mUiKuzB-VFmGD7g). This sheet contains details of the submission, and also a zip file with the necessary files; `get_submissions()` will authenticate against both.

On a regular (at least weekly) basis, download newly submitted items:

```{r get, eval = FALSE, echo=TRUE}
get_submissions()
```

This will create a new folder in the `Submissions` directory of the `articles` repo, and a `DESCRIPTION` file with the meta data.

You should check that this file has been constructed correctly, and that the zip file has unpacked into the top level of the folder. Common missing items are the `Suppl:` line which should list the `.R` files and any data needed to reproduce the paper.

# Check relevance and reproducibility

First check that the paper fits the [scope of the journal](https://journal.r-project.org/#article-types). If not, `reject()` the paper.

Then check that the paper is reproducible by either: (a) compiling the `Rmd` file; or (b) compiling both the LaTeX file and all `.R` files. If there are any problems, request a `resubmission()` from the authors.

# Acknowledge submission and assigning a new article to an EE

For revisions, use `acknowledge_revision()` to send an acknowledgement email to the authors, cc'ing the handling editor.

For new submissions, choose one of the three active EE's (i.e., not the previous EIC). To balance EE load, use `assignments()` to see how many papers each EE is currently handling, and what has been assigned to them in recent months. As far as possible, take account of the relative expertise of each EE so that papers are assigned to the most appropriate person. Use `acknowledge_submission(editor = "XX")`. to send an acknowledgement email to the authors, cc'ing the chosen handling editor.

# Check progress of all papers

The function `report()` shows the status of all current papers being handled, by author and by status. Anything with 3 stars needs urgent attention.

# New executive editor

The EIC is responsible to navigating the search for a new EE. The search for a new editorial board member should begin in the middle of the year, giving enough time to find a replacement before the outgoing member leaves at the end of December. Nominations for the new member are discussed by the editorial board first, and then preferably also with the advisory committee so that they can provide historical perspective.

Once a shortlist of candidates is created, but before approaching anyone on the list, it should be sent to the R Foundation Board members (R-foundation-board@r-project.org) for their feedback and approval. People on the list are then approached by the EIC or an EE, and once someone agrees, the EIC informs the members of the R Foundation (R-foundation-members@r-project.org). The new editor is formally appointed by the R Foundation president.

Once a new EE is found, the following steps take place:

- The new EE is added to the Editors team at `https://github.com/rjournal` and to the `rjournal.slack.com` organisation.
- The new EE is added to the `editors.csv` file in the `inst` folder of the `rj` package.

The EIC may take on outstanding handling editor duties of the outgoing member.

# New associate editors

The EIC is responsible for recruiting new AEs, after discussion with all editors. Ideally, keywords of submissions over the past year are summarised, and compared with keywords of current AEs. New AEs should be recruited for topics where there is the most need.

Once a new AE is appointed, the following steps take place:

- The EIC informs the other EEs.
- The EIC sets up a GitHub repo of the form ae-articles-xx where xx is the initials of the AE
- The new AE is added to the `associate-editors.csv` file in the `inst` folder of the `rj` package.

# Handover to new Editor-in-Chief

* Update the [R Journal Wikipedia page](https://en.wikipedia.org/wiki/The_R_Journal) with the change of editors.
* Ask Martin Maechler to forward r-journal@r-project.org to the new EIC's preferred email address.
* Grant permission to the new EIC to access the [Google sheet](https://docs.google.com/spreadsheets/d/15Tem82ikjHhNVccZGrHFVdzVjoxAmwaiNwaBE95wu6k/edit#gid=1671813370), and [Google drive](https://drive.google.com/drive/folders/0B2aFIue-Ar8dfmdVdHlpTnRndWhxQnNtR0o2YnoycDU1cVVzRlRaM0VSQWU1YUlnZlkyZEk?resourcekey=0-pzsHNE5mUiKuzB-VFmGD7g) used for submissions.
* Hand over the `rjournal.submission` Google account. This account owns the google forms and spreadsheets used by the editors.

# Archives

[Not done since 2019. Is it needed?]

In the `articles` repo, the `Proofs` folder contains all the supporting files of Accepted articles. The `Rejected` folder contains all of the supporting files for rejected submisisons.

From time to time, papers with dates older than two years should be moved to the `archive` repo, to make the `articles` repo smaller. Recommend that this is done at the hand-over of the EiC role at the end of each year.
